\documentclass{article}

\title{Inventories}
\author{Toby Banks}
\date{November 1st, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%%
\section{Inventories}
\begin{itemize}
  \item Explain IFRS and U.S. GAAP rules for determining inventory cost, including which costs are capitalized and methods for allocating costs between costs of goods sold and inventory. 
  \item Discuss how inventories are reported on the financial statements and how the lower of cost or net realizable value is used and applied. 
  \item Compute ending inventory balances and cost of goods sold using the FIFO, weighted average cost, and LIFO methods to account for product inventory and explain the relationship among and the usefulness of inventory and costs of goods  sold data provided by the FIFO, weighted average costs, and LIFO methods when pries are 1) stable, 2) decreasing, or 4) increasing.
  \item Discuss and calculate ratios for useful for evaluating inventory management.
  \item Analyze the financial statements of companies using different inventory accounting methods by comparing and describing the effect of different methods on cost of goods sold, inventory balances, and other financial statement items. 
  \item Compute and describe the effects of the choice of inventory method on profitability, liquidity, activity and solvency ratios.
  \item Calculate adjustments to reported financial statements related to inventory assumptions to aid in comparing and evaluating companies.
  \item Discuss the reasons that a LIFO reserve might rise or decline during a given period and discuss the implications for financial analysis.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Inventories}
\begin{enumerate}
  \item 
\end{enumerate}

\end{document}

