\documentclass{article}
\usepackage{amsmath}

\title{Sampling and Estimation}
\author{Toby Banks \\ HB-Studios}
\date{August 29th, 2009}

\begin{document}

\section{Definitions}
\textbf{Null Hypothesis}
A null hypothesis is the hypothesis to be tested.

\textbf{Alternative Hypothesis}
Is the hypothesis that is accepted if the null hypothesis is rejected. 

\textbf{Test Statistic}
A test statistic is a quantity, calculated based on sample, whose value is the basis for deciding whether or not to reject the null hypothesis. 
%\begin{displaymath}
\[
\text{Test Statistic} = {\text{Sample Statistic} }
%- \text{Value of the Population Parameter under H_0}} \over{\text{Standard Error of the Sample Statistic}}
\]
%\end{displaymath}

\textbf{Type I Error}
A Type I Error occurs when we reject a true null hypothesis. 

\textbf{Type II Error}
A Type II Error occurs when we do NOT reject a false null hypothesis. 

\textbf{Power of a Test}
Power of a test is the probability of correctly rejecting the null hypothesis. 

\textbf{Rejection Point (Critical Value)}
A rejection point (critical value) for a test statistic is a value with which the computed test statistic is compared to decide whether to reject or not reject the null hypothesis.



\section{Deciding which statistic test to use}
We make mathematical decisions based on whether we have a normally distributed population, a large sample size, known or unknown variance.
\subsection{t-statistic}
It's appropriate to use a t-statistic when there is a normally distributed population with unknown variance. Also when the sample size is small there is no alternative (sample size < 30). When variance is not known, a t-statistic is always to be used. Although with sample sizes >= 30 it is possible to use a z-statistic the t-statistic is a more conservative estimate. 

\subsection{z-statistic} 
It's appropriate to use a z-statistic when there is a normally distributed population with known variance. You do not need to calculate degrees of freedom when using a z-statistic because we can use variance to calculate the standard error.\\

\textbf{I must remember the following rejection points as they are not a part of a table.}\\

\begin{enumerate}
  \item Significance level of \(\alpha = 0.10.\)
  \begin{enumerate}
      \item \(H_0:\theta = \theta_0\) versus \(H_\emph{a}:\theta \neq \theta_0\). The rejection points are \(z_{0.05} = 1.645\) and \(-z_{0.05} = -1.645\).
      \item \(H_0:\theta \leq \theta_0\) versus \(H_\emph{a}:\theta > \theta_0\). The rejection points are \(z_{0.10} = 1.28\).
      \item \(H_0:\theta \geq \theta_0\) versus \(H_\emph{a}:\theta < \theta_0\). The rejection points are \(-z_{0.10} = -1.28\).
  \end{enumerate}
  \item Significance level of \(\alpha = 0.05\).
  \begin{enumerate}
      \item \(H_0:\theta = \theta_0\) versus \(H_\emph{a}:\theta \neq \theta_0\). The rejection points are \(z_{0.025} = 1.96\) and \(-z_{0.025} = -1.96\).
      \item \(H_0:\theta \leq \theta_0\) versus \(H_\emph{a}:\theta > \theta_0\). The rejection points are \(z_{0.05} = 1.645\).
      \item \(H_0:\theta \geq \theta_0\) versus \(H_\emph{a}:\theta < \theta_0\). The rejection points are \(-z_{0.05} = -1.645\).
  \end{enumerate}
  \item Significance level of \(\alpha = 0.01\).
  \begin{enumerate}
      \item \(H_0:\theta = \theta_0\) versus \(H_\emph{a}:\theta \neq \theta_0\). The rejection points are \(z_{0.005} = 2.575\) and \(-z_{0.025} = -2.575\).
      \item \(H_0:\theta \leq \theta_0\) versus \(H_\emph{a}:\theta > \theta_0\). The rejection points are \(z_{0.01} = 2.33\).
      \item \(H_0:\theta \geq \theta_0\) versus \(H_\emph{a}:\theta < \theta_0\). The rejection points are \(-z_{0.01} = -2.33\).
  \end{enumerate}
\end{enumerate}
\textbf{NOTE:} That if the first term in a question is like \(H_0:\mu \leq x\) then the answer will be a positive number.\\

\subsection{chi-square test}
It's appropriate to use a chi-square test when there is a single normally distributed population and we know variance. 
\[
{\chi^2} = \frac{\left(n - 1\right) s^2}{\sigma_0^2}
\]
\textit{Rejection Points}
\begin{enumerate}
  \item "Not equal to". Reject the null hypothesis if the test statistic is greater than the upper \(\alpha / 2\) point (denoted \(\chi_{\alpha / 2}^2\)) or less than the lower \(\alpha / 2\) point (denoted \(\chi_{1 - \alpha / 2}^2\)) of the chi-square distribution with df = n - 1.
  \item "Greater than". Reject the null hypothesis if the test statistic is greater than the upper \(\alpha\) point of the chi-square distribution with df = n - 1.
  \item "Less than". Reject the null hypothesis if the test statistic is less than the lower \(\alpha\) point of the chi-square distribution with df = n - 1.
\end{enumerate}

\end{document}

