\documentclass{article}

\title{Long-Term Liabilities and Leases}
\author{Toby Banks}
\date{November 11th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%%
\section{Long-Term Liabilities and Leases}
\begin{itemize}
  \item Compute the effects of debt issuance and amortization of bond discounts and premiums on financial statements and ratios.
  \item Explain the role of debt conenants in protecting creditors by restricting a company's ability to invest, pay dividends, or make other operating and strategic decisions.
  \item Describe the presentation of, and disclosures relating to, financing liabilities.
  \item Determine the effects of changing interest rates on the market value of debt and on financial statements and ratios. 
  \item Describe two types of debt with equity features (convertible debt and debt with warrants) and calculate the effect of issuance of such instruments on a company's debt ratios. 
  \item Discuss the motivations for leasing assets instead of purchasing them and the incentives for reporting the leases as operating leases rather than finance leases.
  \item Determine the effects of finance and operating leases on the financial statements and ratios of the lessess and lessors.
  \item Distinguish between a sales-type lease and a direct financing lease, and determine the effects on the financial statements and ratios of the lessors.
  \item Describe the types and economic consequences of off-balance sheet financing and determine how take-or-pay contracts, throughput arrangements, and the sale of receivables affect financial statements and selected financial ratios. 
\end{itemize}

%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Long-Term Liabilities and Leases}
\begin{enumerate}
  \item 
\end{enumerate}

\end{document}

