\documentclass{article}

\title{Understanding the Cash Flow Statement}
\author{Toby Banks}
\date{November 1st, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%%
\section{Understanding the Cash Flow Statement}
\begin{itemize}
  \item Compare and contrast cash flows from operating, investing, and financing activities and classify cash flow items as relating to one of these three categories, given a description of the items.
  \item Describe how noncash investing and financing activities are reported.
  \item Compare and contrast the key differences in cash flow statements prepared under international financial reporting standards and U.S. GAAP.
  \item Demonstrate the difference between the direct and indirect methods of presenting cash from operating activities and explain the arguments in favor of each.
  \item Demonstrate the steps in the preparation of direct and indirect cash flow statements, including how cash flows can be computed using income statement and balance sheet data.
  \item Describe the process of converting a cash flow statement from the indirect to direct method of presentation.
  \item Analyze and interpret a cash flow statment using both total currency amounts and common-size cash flow statements.
  \item Explain and calculate free cash flow to the firm, free cash flow to equity, and other cash flow ratios.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Understanding the Cash Flow Statement}
\begin{enumerate}
  \item 
\end{enumerate}

\end{document}

