\title{Monopolistic Competition and Oligopoly}
\author{Toby Banks}
\date{September 20th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% What is Monopolistic Competition? %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What is Monopolistic Competition?}
Monopolistic competition is a market structure in which:
\begin{itemize}
  \item A large number of firms compete
  \item Each firm produces a differentiated product
  \item Firms compete on product quality, price, and marketing
  \item Firms are free to enter and exit
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Price and Output in Monopolistic Competition  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Price and Output in Monopolistic Competition} 
Profit is maximized where marginal revenue equals marginal cost. \\
In the short run, a firm in monopolistic competition makes its output and price decisions the same way that a monopoly does. \\
Profit maximizing is equivilant to loss minimizing if the circumstances are right. \\
The long-term trend for economic profit is to trend towards zero economic profit. \\
The key differences between monopolistic competition and perfect competition are:
\begin{itemize}
  \item Excess Capacity (firms could sell more by cutting prices but by doing so they will incur a economic loss).
  \item Markup (The amount a price exceeds marginal cost).
\end{itemize}

\hfill \\
Is monopolistic competition efficient? Well not really in the sense that perfect competition is efficient. It's kind of the wrong question. \\
The markup that drives a gap between price and marginal cost in monopolistic competition arises from product differentiation. \\
Product variety is both valued and costly. The efficient degree of product variety is one for which the marginal social benefit of product variety equals its marginal social cost. So compared to the alternative - product uniformity - monopolistic competition might be efficient. \\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Product Development and Marketing %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Product Development and Marketing}
A firm trys to maintain economic profit by seeking out new products that will provide it with a competitive edge. \\
Firms can also use advertising as a way of differentiating their product form the competition. \\
The marginal social benefit of an innovation is the increase in price that consumers are willing to pay for it. \\

 
\end{document}

