\documentclass{article}

\title{Output and Costs - Questions}
\author{Toby Banks}
\date{September 19th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%
\section{Output and Costs - Learning Outcomes}
\begin{itemize}
  \item Differentiate between short-run and long-run decision time frames.
  \item Describe and explain the relations among total product of labor, marginal product of labor, and average product of labor, and describe increasing and decreasing marginal returns.
  \item Distinguish among total cost (including both fixed cost and variable cost), marginal cost, and average cost, and explain the relations among the various cost curves.
  \item Explain the company's production function, its properties of diminishing returns and diminishing marginal product of capital, the relation between short-run and long -run costs, and how economies and diseconomies of scale affect long-run costs. 
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Output and Costs - Questions}
\begin{enumerate}
  \item Describe the shape of the total product curve.
  \item Describe the shape of the average product curve. How does it relate to the marginal product curve.
  \item Define the following: Total Product, Marginal Product, Average Product, Increasing Marginal Returns, Diminishing Marginal Returns, Total Cost, Marginal Cost, Average Cost, Total Fixed Cost, Total Variable Cost, Average Fixed/Variable/Total Cost.
  \item What is the "law of diminishing returns"?
  \item Why is the Average Total Cost curve U-Shaped?
  \item The position of a firm's short-run cost curve depends on two factors. Describe these factors.
  \item Can you describe the short-run costs of four different plants via a long-run average cost curve?
  \item What are economies of scale, diseconomies of scale, constant returns to scale, and minimum efficient scale?
\end{enumerate}

\end{document}

