\documentclass{article}

\title{Perfect Competition - Questions}
\author{Toby Banks}
\date{September 20th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%
\section{Perfect Competition - Learning Outcomes}
\begin{itemize}
  \item Describe the characteristics of perfect competition, explain why companies in a perfectly competitive market are price takers, and differentiate between market and company demand curves.
  \item Determine the profit maximizing (loss minimizing) output for a perfectly competitive company and explain marginal cost, marginal revenue, and economic profit and loss.
  \item Describe a perfectly competitive company's short-run supply curve and explain the impact of changes in demand, entry and exit of companies, and changes in plant size on the long-run equilibrium.
  \item Discuss how permanent change in demand or changes in technology affect price, output, and economic growth.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Perfect Competition - Questions}
\begin{enumerate}
  \item What traits define a industry as being perfectly competitive?
  \item What two factors lead to perfect competition?
  \item What is the difference between Marginal Revenue (MR), Total Revenue (TR), Total Cost (TC), and Economic Profit?
  \item What decisions can be made in the short-run and long-run in a perfectly competitive industry?
  \item Draw a pretend graph that details the relationship between Total Cost and Total Revenue in a perfectly competitive industry. Where do economic losses and profits occur?
  \item How can we use marginal analysis to determine how much of a good to produce?
  \item What is a shutdown point? Draw a graph detailing where the shutdown point occurs. Hint, marginal cost, marginal revenue and average variable cost are included.
  \item Draw a pretend graph that details the relationship between Marginal Cost and Marginal Revenue. Where is the profit maximization point?
  \item What is a short-run supply curve?
  \item What effect does the entry of new firms into a perfectly competitive market have?
  \item What effect does the exit of old firms from a perfectly competitive market have?
  \item What effect does increasing the size of a plant have in a perfectly competitve market?
  \item What effect does decreasing the size of a plant have in a perfectly competitve market?
  \item What effect does both a permenant increase and decrease in demand have on a perfectly competitive market?
  \item What are external economies and external diseconomies? Can you draw their effects on a graph?
\end{enumerate}

\end{document}

