\documentclass{article}

\title{Money, the Price Level, and Inflation}
\author{Toby Banks}
\date{September 27th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%
\section{Money, the Price Level, and Inflation}
\begin{itemize}
  \item Explain the functions of money.
  \item Describe the components of the M1 and M2 measures of money and discuss why checks and credit cards are not counted as money.
  \item Describe the economic functions of and differentiate among the various depository institutions and explain the impact of financial regulation, deregulation, and innovation.
  \item Explain the goals of the U.S. Federal Reserve (Fed) in conducting monetary policy and how the Fed uses its policy tools to control the quantity of money, and describe the assets and liabilities on the Fed's balance sheet. 
  \item Discus the creation of money, including the role played by excess reserves, and calculate the amount of loans a bank can generate, given new deposits.
  \item Describe the monetary base and explain the relation among the monetary base, the money multiplier, and the quantity of money.
  \item Explain the factors that influence the demand for money and describe the demand for money curve, including the effects of changes in real GDP and financial innovation.
  \item Explain interest rate determination and the short-run and long-run effects of money on real GDP.
  \item Discuss the quantity theory of money and its relation to aggregate supply and aggregate demand.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Money, the Price Level, and Inflation}
\begin{enumerate}
  \item What are the four functions of money?
  \item What are the three types of depository institutes that make up the nation's money?
  \item What three types of assets do commercial banks have?
  \item What are the four main servies that people are willing to pay for when they give their money to a depository institution?
  \item What are the three main influences on financial innovation?
  \item Describe the structure of the Federal Reserve. What are its main three policy tools? What are the three categories of assets the Fed has? What are the two categories of liabilities?
  \item What are the three factors that limit the quantity of deposits that the banking system can create?
  \item What is the currency drain ratio?
  \item Describe the desired reserve ratio?
  \item What are excess reserves?
  \item Describe the nine steps in the money creation process?
  \item What is the money multiplier?
  \item What are the four main influences on the quantity of oney that people plan to hold?
  \item What is the definition of the demand for money?
  \item Define velocity of circulation.
\end{enumerate}

\end{document}

