\documentclass{article}

\title{Money, the Price Level, and Inflation}
\author{Toby Banks}
\date{September 28th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% What is Money? %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What is Money?}
Money is any commodity or token that is generally acceptable as a means of payment. If functions as 
\begin{itemize}
  \item A medium of exchange
  \item A unit of account
  \item A store of value
\end{itemize}
\hfill \\

Today, money consists of currency and deposits at banks and other depository institutions. \\
\hfill \\
M1 consists of currency and traveler's checks plus checking deposits owned by individuals and businesses. M1 \textbf{does not} include currency held by banks or currency and checking deposits owned by the U.S> government. \\
\hfill \\
M2 consists of M1 plus time deposits, savings deposits, and money market mutual funds and other deposits. M2 is almost five times as large as M1 in the United States. \\
\hfill \\


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Depository Institutions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Depository Institutions}
Commercial banks, Thrift institutions (such as SLs, savings banks, and credit unions), as well as  money market mutual funds are depository institutions whose deposts are money. \\
\hfill \\
A commercial bank has three types of assets:
\begin{itemize}
  \item Liquid assets (such as Treasury Bills and commercial bills).
  \item Investment securitites (longer-term U.S. government bonds and other bonds).
  \item Loans (mortgages, small business loans, personal loans, credit card debt).
\end{itemize}
\hfill \\

Depository institutuions procide four main economic services: 
\begin{itemize}
  \item Create liquidity
  \item Minimize the cost of obtaining funds
  \item Minimize the cost of monitoring borrowers
  \item Pooling risk
\end{itemize}
\hfill \\

The three main influences on financial innovation are:
\begin{itemize}
  \item Economic Environment
  \item Technology
  \item Regulation
\end{itemize}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% The Federal Reserve System %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Federal Reserve System}
The Fereral Reserve System is the central bank of the United State. \\
\hfill \\
The Fed influences the quantity of money by setting the required reserve ratio, the discount rate, and by conducting open market operations. \\
\hfill \\
The key elements of the Federal Reserve Bank are:
\begin{itemize}
  \item The Board of Governors (Seven members appointed to 14 year terms by the President, one member is the chairman for four year terms).
  \item The regional Federal Reserve Banks (There are 12 and the New York one is special because it implements the policy decisions of the Federal Open Market Committee).
  \item The Federal Open Market Committee 
\end{itemize}
\hfill \\

The \textbf{Federal Open Market Committee} is the main policymaking organ of the Federal Reserve System. The following voting memvers meet approximately every six weeks to review the state of the economy.
\begin{itemize}
  \item The chairman and other six members of the Board of Governors
  \item The president of the Federal Reserve Bank of New York
  \item The presidents of the other regional Federal Resere banks (of whom on a yearly rotating basisi only four vote)
\end{itemize}
\hfill \\

The Fed has three main policy tools at its disposal:
\begin{itemize}
  \item Required Reserve Ratios
  \item Discount Rate
  \item Open Market Operations (the purchase and sale of government securities, U.S. T-Bills and bonds to the open market)
\end{itemize}
\hfill \\

The Fed has the following assets:
\begin{enumerate}
  \item Gold and foreign exchange
  \item U.S. Government Securities
  \item Loans to banks
\end{enumerate}
\hfill \\

The Fed has the following liabilites:
\begin{enumerate}
  \item Federal Reserve Notes
  \item Banks' deposits
\end{enumerate}
\hfill \\

The \textbf{monetary base} is the sum of Federal Reserve notes, coins, and banks' deposits at the Fed. \\


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% How Banks Create Money %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How Banks Create Money}
Banks create money by making loans. \\
\hfill \\
The total quantity of money that can be created depends on 
\begin{itemize}
  \item Monetary base
  \item The desired reserve ratio 
  \item The currency drain ratio
\end{itemize}
\hfill \\

A bank's actual \textbf{reserves} consist of the notes and coins in its vaults and its deposit at the Federal Reserve. \\
\hfill \\
The \textbf{reserve ratio} is hte fraction of a bank's total deposits that are held in reserves. \\
\hfill \\
The \textbf{desired reserve ratio} is the ratio of reserves to bank deposits that a bank wants to hold. This ratio exceed the required reserve ratio by an amount htat the banks determine to be prudent on the basis of their daily business requirements. \\
\hfill \\
A bank's \textbf{excess reserves} are its actual reserves minus its desired reserves. Whenever the banking system as a whole has excess reseres, the banks are able to create money. \\
\hfill \\

The following sequence repeats until the bank has no excess reserves:
\begin{enumerate}
  \item Banks have excess reserves.
  \item Banks lend excess reserves.
  \item Bank deposits increase.
  \item The quantity of money increases.
  \item New money is used to make payments.
  \item Some of the new money remains on deposit.
  \item Some of the new money is in a \emph{currency drain}.
  \item Desired reserves increase because deposits have increased.
  \item Excess reserves decrease bu remain positive.
\end{enumerate}
\hfill \\

The \textbf{money multiplier} is the ratio of the change in the quantity of money to the change in the monetary base. The equation is \(A / (1 - L) \). \\
\hfill \\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% The Market for Money %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Market for Money}
The quantity of money demanded is the amount of money that people plan to hold. \\
\hfill \\
The quantity of real money equals the quantity of nominal money divided by the price level. \\
\hfill \\
The quantity of real money demanded depends on 
\begin{itemize}
  \item The price level
  \item Nominal interest rate
  \item Real GDP
  \item Financial innovation 
\end{itemize}
\hfill \\

A rise in the nominal interest rate brings a decrease in the quantity of real money demanded. \\
\hfill \\
In the short run, the Fed sets the quantity of money to hit a target nominal interest rate. \\
\hfill \\
In the long run, the loanable funs market determines the interest rate and the money market equilibrium determines the price level. \\
\hfill \\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% The Quantity Theory of Money %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Quantity Theory of Money}
The quantity theory of money is the proposition that moeny growth and inflation move up and down together in the long run. \\
\hfill \\
The U.S. and international evidence is consistent with the quantity theory, on the average. \\
\hfill \\


\end{document}

