\documentclass{article}

\title{Income Taxes}
\author{Toby Banks}
\date{November 10th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%%
\section{Income Taxes}
\begin{itemize}
  \item Explain the differences between accounting profit and taxable income, and define key terms, including deferred tax assets, deferred tax liabilities, valuation allowance, taxes payable, and income tax expense. 
  \item Explain how deferred tax liabilities and assets are created and the factors that determine how a company's deferred tax liabilities and assets should be treated.
  \item Determine the tax base of a company's assets and liabilities.
  \item Calculate income tax expense, income taxes payable, deferred tax assets and deferred tax liabilities, and calculate and interpret the adjustment to the financial statements related to a change in the income tax rate.
  \item Evaluate the impact of tax rate changes on a company's financial statements and ratios
  \item Distinguish between temporary and permanent items in pre-tax financial income and taxable income.
  \item Discuss the valuation allowance for deferred tax assets - when it is required and what impact it has on financial statements.
  \item Compare and contrast a company's deferred tax items.
  \item Analyze disclosures relating to deferred tax items and the effective tax rate reconciliation, and discuss how information included in these disclosures affects a company's financial statements and financial ratios.
  \item Identify the key provisions of and differences between income tax accounting under U.S. GAAP and IFRS.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Income Taxes}
\begin{enumerate}
  \item 
\end{enumerate}

\end{document}

