\documentclass{article}

\title{U.S. Inflation, Unemployment, and Business Cycles}
\author{Toby Banks}
\date{September 27th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%
\section{U.S. Inflation, Unemployment, and Business Cycles}
\begin{itemize}
  \item Differentiate between inflation and the price level.
  \item Describe and distinguish among the factors resulting in demand-pull and cost-push inflation and describe the evolution of demand-pull and cost-push inflationary processes.
  \item Explain the costs of anticipated inflation.
  \item Explain the relation among inflation, nominal interest rates, and the demand and supply of money.
  \item Explain the impact of inflation on unemployment and describe the short-run and long-run Phillips curve, including the effect of changes in the natural rate of unemployment.
  \item Explain how economic growth, inflation, and unemployment affect the business cycle.
  \item Describe mainstream business cycle theory and real business cycle (RBC) theory and distinguish between them, including the role of productivity changes. 
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{U.S. Inflation, Unemployment, and Business Cycles}
\begin{enumerate}
  \item Describe in lots of detail Demand-Pull inflation.
  \item Describe in lots of detail Cost-Push inflation.
  \item Describe the effect of expected inflation.
  \item Describe the short-run Phillips curve.
  \item Describe the long-run Phillips curve.
  \item Provide a brief overview of the following mainstream business cycle theories:
  \begin{enumerate}
    \item Keynesian Cycle Theory.
    \item Monetarist Cycle Theory.
    \item New Classical Cycle Theory.
    \item New Keynesian Cycle Theory.
  \end{enumerate}
  \item Describe in great detail the real business cycle theory.
  \item What are the criticisms and defenses of RBC theory?
\end{enumerate}

\end{document}

