\documentclass{article}

\title{U.S. Inflation, Unemployment, and Business Cycles}
\author{Toby Banks}
\date{September 28th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% The Evolving U.S. Economy %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Evolving U.S. Economy}
Potential GDP Grows. \\
\hfill \\
Inflation persists because aggregate demand grows faster than potential GDP. \\
\hfill \\
Business cycles occur because potential GDP and aggregate demand change at an uneven pace. \\
\hfill \\
Demand-pull inflation is triggered by an increase in aggregate demand and fueled by ongoing money growth. Real GDP cycles above full employment. \\
\hfill \\
Cost-push inflation is triggered by an increase in the money wage rate or raw material prices and is fueled by ongoing money growth. Real GDP cycles below full employment in a stagflation. \\
\hfill \\
When inflation is expected correctly, real GDP remains at potential GDP. \\
\hfill \\
The short-run Phillips curve shows the trade off between inflation and unemployment when the expected inflation rate and the natural unemployment rate are constant. \\
\hfill \\
The long-run Phillips curve, which is vertical, shows that when the actual inflation rate equals the expected inflation rate, the unemployment rate equals the natural unemployment rate. \\
\hfill \\
The mainstream business cycle theory explains business cycles as fluctuations of real GDP around potential GDP and as arising from a steady expansion of potential GDP combined with an expansion of aggregate demand at a fluctuating rate. \\
\hfill \\
Real business cycle theory explains business cycles as fluctuations of potential GDP, which arise from fluctuations in the influence of technological change on productivity growth.
\hfill \\

\end{document}

