\documentclass{article}

\title{Financial Statement Analysis: An Introduction - Questions}
\author{Toby Banks}
\date{October 12th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%%
\section{Financial Statement Analysis: An Introduction - Learning Outcomes}
\begin{itemize}
  \item Discuss the roles of financial reporting and financial statement analysis.
  \item Discuss the role of key financial statements (income statement, balance sheet, statement of cash flows, and statement of changes in owners' equity) in evaluating a company's performance and financial position.
  \item Discuss the importance of financial statement notes and supplementary information, including disclosures of accounting methods, estimates, and assumptions, and managements' discussion and analysis.
  \item Discuss the objective of audits of financial statements, the types of audit reports, and the importance of effective internal controls.
  \item Identify and explain information sources of other than the annual financial statements and supplementary information that analysts use in financial statement analysis.
  \item Describe the steps in the financial statement analysis framework. 
\end{itemize}


%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Financial Statement Analysis: An Introduction - Questions}
\begin{enumerate}
  \item What is the difference between profit and cash flow?
  \item Give four examples of economic decisions that may rest on financial statement analysis?
  \item What is the basic equation of a income statement?
  \item What types of questions would you look to answer with a income statement?
  \item What are two alternative names for the income statement?
  \item What are two alternative names for the balance sheet?
  \item What is the basic equation of a balance sheet?
  \item What types of questions would you seek to answer with a balance sheet?
  \item Define financial flexibility, operating activities, investing activities, and financing activities.
  \item What is the Statement of Change in Owners' Equity?
  \item What are some examples of footnotes?
  \item What do the following acronyms stand for: GAAP, I.F.R.S., and F.A.S.B.
  \item What is the purpose of the MD\&A statement?
  \item What are the four paragraphs in a auditors report? What does each contain.
  \item What are the four different auditors opinions that might be expressed? What do they each mean?
  \item In regards to internal controls, what four requirements are placed on firms by regulators?
  \item What five things must a management report on internal control contain according to the Sarbanes-Oxley Act?



  \item What are the six individual phases of Financial Statement Analysis Framework?
  \item What are the three inputs and four outputs to the "Articulate the purpose and context of the analysis" phase?
  \item What are the three inputs and three outputs to the "Collect Data" phase?
  \item What are the one input and four outputs to the "Process Data" phase?
  \item What are the one input and one output to the "Analyze/Interpret the processed data" phase?
  \item What are the two inputs and two outputs to the "Develop and Communicate conclusions and recommendations" phase?
  \item What are the one input and one output to the "Follow Up" phase?
\end{enumerate}

\end{document}

