\documentclass{article}

\title{Aggregate Supply and Aggregate Demand}
\author{Toby Banks}
\date{September 27th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%
\section{Aggregate Supply and Aggregate Demand}
\begin{itemize}
  \item Explain the factors that influence real GDP and long-run and short-run aggregate supply. Explain movement along the long-run and short-run aggregate supply curves (LAS and SAS). Discuss the reasons for changes in potential GDP and aggregate supply.
  \item Explain the components of and the factors that affect real GDP demand. Describe the aggregate demand curve and why it slopes downward. Explain the factors that can change aggregate demand.
  \item Differentiate between short-run and long-run macroeconomic equilibrium and explain how economic growth, inflation, and changes in aggregate demand and supply influence the macroeconomic equilibrium. 
  \item Compare and contrast the classical, Keynesian, and monetarist shcools of macroeconomics.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Aggregate Supply and Aggregate Demand}
\begin{enumerate}
  \item Differentiate between the macroeconomic long and short-run.
  \item Differentiate between the long-run and short-run aggregate supply.
  \item What three events can cause a increase in potential GDP (aka. long-run aggregate supply)?
  \item What is the quantity of real GDP demanded? What is a equation to represent it?
  \item What is aggregate demand?
  \item What two effects cause the aggregate demand curve to slop downward?
  \item What three factors influence buying plans (changes in aggregate demand) other than the price level?
  \item What is the difference between fiscal policy and monetary policy. 
  \item Define disposable income. 
  \item Contrast short-run macroeconomic equilibrium and long-run macroeconomic equilibrium.
  \item Use three graphs to describe the business cycle. Detail recessionary gap, inflationary gap, and full employment in the three graphs.
  \item What is stagflation?
  \item Detail how classical, Keynesian, and monetarist economists view aggregate demand fluctuations, the response of aggregate supply, and the policy to deal with it.
  \item What is a new Keynesian economists view?
  \item What is a new classical economists view?
\end{enumerate}

\end{document}

