\documentclass{article}

\author{Toby Banks}
\date{Dec 11th, 2009}

\begin{document}

\section{Cost of Capital - LOS}
\begin{itemize}
  \item Calculate and interpret the weighted average cost of capital (WACC) of a company.
  \item Describe how taxes affect the cost of capital from different capital sources.
  \item Describe alternative methods of calculating the weights used in the WACC, including the use of the company's target capital structure.
  \item Explain how the marginal cost of capital and the investment opportunity schedule are used to determine the optimal capital budget.
  \item Explain the marginal cost of capital's role in determining the net resent value of a project.
  \item Calculate and interpret the cost of fixed rate debt capital using the yield-to-maturity approach and the debt-rating approach.
  \item Calculate and interpret the cost of non-callable, nonconvertible preferred stock.
  \item Calculate and interpret the cost of equity capital using the capital asset pricing model approach, the dividend discount model approach, and the bond-yield-plus risk-premium approach.
  \item Calculate and interpret the beta and cost of capital for a project.
  \item Explain the country equity risk premium in the estimation of the cost of equity for a company located in a developing market.
  \item Describe the marginal cost of capital schedule, explain why it may be upward-sloping with respect to additional capital, and calculate and interpret its break-points.
  \item Explain and demonstrate the correct treatment of flotation costs.
\end{itemize}

\section{Cost of Capital - Questions}
\begin{enumerate}
  \item What is the formula for the weighted average cost of capital?
  \item Given the D/E ratio calculate the weight for debt and equity?
  \item What are three approaches that an analyst might use to estimate the target capital structure of a company?
  \item Describe the investment opportunity schedule. Draw a diagram of it.
  \item Calculate the cost of debt using the yield to maturity approach. NOTE: Their are calculator tricks to know.
  \item What is the debt-rating approach to estimating the cost of debt?
  \item What are four issues in estimating the cost of debt?
  \item What is the formula to calculate the value of preferred stock? How can this be rearranged to calculate the cost of preferred stock?
  \item What is the formula for the CAPM? What does it give you?
  \item What are the three formulas for the Dividend Discount Model Approach? 
  \item What is the formula for the sustainable growth rate?
  \item What is the formula for the Bond Yield plus Risk Premium Approach?
  \item How can you estimate a companies stock beta by using a market regression model?
  \item Beta estimates are sensitive to the method of estimation and data used. Give five examples of issues surrounding this.
  \item What is the formula to calculate the asset beta?
  \item What is the formula to calculate the equity beta?
  \item What are the four steps in estimating a beta using the pure-play method?
  \item What is the formula to calculate the beta of a comparable company?
  \item Assuming that we have the beta of a comparable company, how do we calculate the beta for our project?
  \item Calculate the before-tax marginal cost of debt. What is the formula?
  \item 

\end{enumerate}

\end{document}

