\documentclass{article}

\title{Organizing Production - Questions}
\author{Toby Banks}
\date{September 19th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%
\section{Organizing Production - Learning Outcomes}
\begin{itemize}
  \item Explain the types of opportunity cost and their relation to economic profit, and calculate economic profit.
  \item Discuss a company's constraints and their impact on the achieve-ability of maximum profit.
  \item Differentiate between technological efficiency and economic efficiency and calculate economic efficiency of various companies under different scenarios.
  \item Explain command system and incentive systems to organize production, the principal-agent problem, and measures a firm uses to reduce the principal-agent problem.
  \item Describe the different types of business organization and the advantages and disadvantages of each.
  \item Calculate and interpret the four-firm concentration ratio and the Herfindahl-Hirschman Index and discuss the limitations of concentration measures.  
  \item Explain why companies are often more efficient than markets in coordinating economic activity.
\end{itemize}


%%%%%%%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Organizing Production - Questions}
\begin{enumerate}
  \item Define the following: explicit costs, implicit costs, economic depreciation, normal profit, economic profit, technology.
  \item What are the five basic decisions a firm must make when trying to maximize profit?
  \item What are the three features of a firms environment that limit the maximum profit that it can make?
  \item What is production efficiency? What two components make it up?
  \item Firms use both command and incentive systems to organize production. Discuss both.
  \item What is the principal-agent problem? What are three ways of attempting to deal with it?
  \item What are the three main types of business organizations? Describe the pros and cons of each.
  \item What are the four market types that economists identify? Describe each.
  \item What are the two measure of concentration. Detail the threshold levels for each as they relate to the four market types.
  \item What are the three main limitations of using only concentration measures as determinants of market structure?
  \item What are the advantages and disadvantages of using markets vs. firms for coordinating decisions? What are the major benefits firms have over markets? 
\end{enumerate}

\end{document}

