\documentclass{article}
\usepackage{amsmath}

\title{Organizing Production}
\author{Toby Banks}
\date{September 19th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%% What the Fuck?                   %%%%%%%%%%%%%%%%%%%%%%%%% 
\section{What the fuck?}
...is accounting depreciation? \\


%%%%%%%%%%%%%%%%%%%%%%%%% The Firm and its Economic Problem %%%%%%%%%%%%%%%%%%%%%%%%
\section{The Firm and its Economic Problem}
\textbf{Opportunity cost} of any action is the highest-valued alternative forgone. Opportunity cost includes both explicit costs and implicit costs. \\
\hfill \\
\textbf{Explicit costs} are paid in money. \\
\hfill \\
\textbf{Implicit costs} occur when a firm forgoes an alternative action but does not make a payment. For example when a firm uses its own capital, or uses its owner's time or financial resources. \\
\hfill \\
\textbf{Implicit rental rate of capital} is the cost of using a firms capital, when it could have rented the capital out. \\
\hfill \\
\textbf{Economic depreciation} is the change in the market value of capital over a given period. It is a implicit cost. \\
\hfill \\
\textbf{Interest forgone} is the interest lost on the money used to buy capital or that is allocated to some other purpose. It is a implicit cost. \\
\hfill \\
\textbf{Accounting costs} largely about explicit costs (for example, the cost of machinery, deprecation of value of assets, etc.) \\
\hfill \\
\textbf{Economic costs} include implicit costs, as well as in the case of a single proprietorship entrepreneurial skills and capital. \\
\hfill \\
\textbf{Normal profit} is the return that a entrepreneur sees on his business. It is a opportunity cost as it could have been reinvested in the business. \\
\hfill \\
\textbf{Economic profit} is \emph{Total Revenue - Total Cost (total cost includes opportunity cost)}. \\
\hfill \\
The objective of economic accounting is to maximize economic profit. To do so a firm must make five basic decisions:
\begin{enumerate}
  \item What goods and services to produce and in what quantities?
  \item How to produce - the techniques of production to use.
  \item How to organize and compensate its managers and workers.
  \item How to market and price its products.
  \item What to produce itself and what to buy from other firms.
\end{enumerate}
\hfill \\
There are three features of a firm's environment that limit the maximum profit that it can make:
\begin{enumerate}
  \item Technology
  \item Information
  \item Market
\end{enumerate}


%%%%%%%%%%%%%%%%%%%%%%%%% Technological and Economic Efficiency %%%%%%%%%%%%%%%%%%%%
\section{Technological and Economic Efficiency}
\textbf{Technological efficiency} is a process that uses the least inputs for a given output. \\
\hfill \\
\textbf{Economic efficiency} is a process that uses the least cost for a given output. \\
\hfill \\
\textbf{Production Efficiency} is a process that is the most technological efficient and economic efficient. \\
\hfill \\


%%%%%%%%%%%%%%%%%%%%%%%%% Information and Organization %%%%%%%%%%%%%%%%%%%%
\section{Information and Organization}
\textbf{Command systems} Hierarchies used for communicate information upwards and barking commands back downwards. \\
\hfill \\
\textbf{Incentive systems} Configuring payments, benefits, and other incentives to employees in a attempt to maximize the profits of the firm. \\
\hfill \\
\textbf{The principal-agent problem} is the problem of devising compensation rules that induce an \emph{agent} to act in the best interest of a \emph{principal}. Three ways of attempting to cope with the principal-agent problem are \textbf{ownership, incentive pay, and long-term contracts}. It is important to note that commands do not work for the principal-agent problem because in most firms the shareholders can effectively monitor the managers and the managers can't effectively monitor the employees. \\
\hfill \\

\begin{tabular}{l p{5cm} p{5cm}}
  \textbf{Type of Firm} & \textbf{Pros} & \textbf{Cons} \\
  \hline
  \textbf{Proprietorship} & Easy to set up & Bad decisions not checked by need for consensus \\
                          & Simple decision making & Owner's entire wealth at risk \\
                          & Profits taxed only once as owner's income & Firm dies with owner \\
                          & & Cost of capital and labor is high relative to that of a corporation \\
  \hline
  \textbf{Partnership}    & Easy to set up & Achieving a consensus may be slow and expensive \\
                          & Diversified decision making & Owners' entire wealth at risk \\
                          & Can survive withdrawal of partner & Cost of capital and labor is high relative to that of a corporation \\
                          & Profits taxed only once as owners' incomes & \\
  \hline
  \textbf{Corporation}    & Owners have limited liability & Complex management structure can make decisions slow and expensive \\
                          & Large-scale, low-cost capital available & Retained profits taxed twice: as company profit and as stockholders' capital gains \\
                          & Professional management not restricted by ability of owners & \\
                          & Perpetual life & \\
                          & Long-term labor contracts cut labor costs& \\
  \hline
\end{tabular}

%%%%%%%%%%%%%%%%%%%%%%%%% Markets and the Competitive Environment %%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Markets and the Competitive Environment}

\hfill \\
\textbf{Perfect competition} is where there are many firms selling \emph{identical} products, many buyers, and no restrictions on the entry of new firms into the industry. Good examples are worldwide markets for corn,rice, and other grain crops. \\
\hfill \\
\textbf{Monopolistic competition} is where a large number of firms compete by making similar but slightly different products. \\
\hfill \\
\textbf{Oligopoly} is a market structure in which a small number of firms compete. Good examples are computer software, airplane manufacture, and international air transportation. \\
\hfill \\
\textbf{Monopoly} is where there is only one firm which produces a good or service with no close substitutes and there is a large protective barrier to entry of new firms. For example, the electric utility. \\
\hfill \\
\hfill \\

\begin{tabular}{l p{6cm}}
  \multicolumn{2}{l}{\textbf{Measures of Concentration}} \\
  \hline
  Four-Firm concentration ratio & The percentage of the value of sales accounted for by the four largest firms in the industry. \\
  Herfindahl-Hirschman Index & The squared percentage of market share of he largest 50 firms in a industry. \\
  \hline
\end{tabular}
\hfill \\
\hfill \\


\begin{tabular}{l|c||c}
  \hline
  \textbf{Competition Level} & Four Firm & Herfindahl-Hirschman \\
  \hline
  \textbf{Perfect Competition} & Close to 0 & Less than 100 \\
  \textbf{Monopolistic Competition} & Low & 101 to 999 \\
  \textbf{Oligopoly} & High & More than 1000 \\
  \textbf{Monopoly} & 100 & 10,000 \\
  \hline
\end{tabular}
\hfill \\
\hfill \\

There are three main limitations of using only concentration measures as determinants of market structure and their failure to take proper account of:
\begin{enumerate}
  \item The geographical scope of the market. Concentration measures take a national view of the market, where sometimes a regional or global view is more accurate.
  \item Barriers to entry and firm turnover. Some industries are highly concentrated but have easy entry and an enormous amount of turnover of firms. For example, a restaurant in a small town. Also, an industry might be competitive because of \emph{potential entry}. For example, there might be only a few firms in a market, but many firms can easily enter the market if economic profits are available.
  \item The correspondence between a market and an industry. The Department of Commerce classifies each firm as being in a particular industry. Markets do not always correspond closely with industries for three reasons:
  \begin{enumerate}
    \item Markets are narrower than industries. For example, the pharmaceutical industry, where there are few firms but many different markets for individual products (drugs). 
    \item Most firms make several products. For example, General Electric.
    \item Firms switch from one market to another based on profit opportunities. For examples, newspapers going online.
  \end{enumerate}
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%% Markets and Firms                %%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Markets and Firms}

Markets are good at coordinating decisions. They do so by adjusting prices and making the decisions of buyers and sellers consistent - making the quantity demanded equal to the quantity supplied for each good and service. \\

Firms are often more efficient than markets as coordinators of economic activity because they can achieve:
\begin{itemize}
  \item Lower transaction costs.
  \item Economies of scale.
  \item Economies of scope (use of specialization and often expensive resources to produce a \emph{range of goods and services}.
  \item Economies of team production (individuals in a group specializing in mutually supportive tasks).
\end{itemize}


\end{document}

