\documentclass{article}
\usepackage{amsmath}

\title{Efficiency and Equity}
\author{Toby Banks}
\date{September 15th, 2009}

\begin{document}

\section{Resource Allocation Methods}

The following list contains mechanisms fro allocating scarce resources. No one system works perfectly for all situations. \emph{Can I immediately think of a example of each?} \\
\begin{itemize}
  \item Market Price
  \item Command
  \item Majority Rule
  \item Contest
  \item First-Come, First-Served
  \item Lottery
  \item Personal Characteristics
  \item Force
\end{itemize}

\hfill \\

\section{Demand and Marginal Benefit}
\begin{itemize}
  \item \textbf{Marginal benefit} determines demand, and a demand curve is a marginal benefit curve. \\
  \item \textbf{The market demand curve} is the horizontal sum of the individual demand curves and is also known as the \textbf{marginal social benefit curve}. \\
  \item \textbf{Value} is what people are willing to pay; \textbf{price} is what they must pay. \\
  \item \textbf{Consumer surplus} equals value minus price, summed over the quantity bought. \emph{Can I mentally picture this on a graph?}  \\
\end{itemize}

\section{Supply and Marginal Benefit}
\begin{itemize}
  \item The minimum supply price determines supply, and the supply curve is the \textbf{marginal cost curve}. \\
  \item The \textbf{market supply curve} is the horizontal sum of the individual supply curves and is the \textbf{marginal social cost curve}. \\
  \item \textbf{Opportunity cost} is what producers pay; price is what producers receive. \\
  \item \textbf{Producer surplus} equals price minus opportunity cost, summed over the quantity sold. \\ 
\end{itemize}

\section{Is the Competitive Market Efficient}
\begin{itemize}
  \item In a competitive equilibrium, marginal social benefit equals marginal social cost and resource allocation is efficient. \\
  \item Buyers and sellers acting in their self-interest end up promoting the social interest. \\
  \item Producing less than or more than the efficient quantity creates \textbf{dead weight loss}. \\
  \item The following can result in \emph{inefficiency and dead weight loss}: \\
  \begin{itemize}
    \item Price and Quantity Regulations
    \item Taxes and Subsidies
    \item Externalities
    \item Public Goods and Common Resources 
    \item Monopoly
    \item High Transaction Costs
  \end{itemize}
\end{itemize}

\section{Is the Competitive Market Fair?}
Ideas about fairness can be divided into two groups: \emph{fair results} and \emph{fair rules}. \\
\begin{itemize}
  \item \textbf{Fair results} ideas require income transfers from the rich to the poor. \\
  \item \textbf{Fair rules} ideas require property rights and voluntary exchange. \\
\end{itemize}
 
\end{document}

