\documentclass{article}

\title{Understanding the Balance Sheet}
\author{Toby Banks}
\date{October 23rd, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%%
\section{Understanding the Balance Sheet}
\begin{itemize}
  \item Illustrate and interpret the components of the balance sheet and discuss the uses of the balance sheet in financial analysis.
  \item Describe the various formats of balance sheet presentation.
  \item Explain how assets and liabilities arise from the accrual process. 
  \item Compare and contrast current and non-current assets and liabilities. 
  \item Explain the measurement bases (e.g. historical cost and fair value) of assets and liabilities, including current assets, current liabilities, tangible assets, and intangible assets.
  \item Demonstrate the appropriate classification and related accounting treatments for marketable and non-marketable financial instruments held as assets or owed by the company as liabilities.
  \item List and explain the components of owners' equity.
  \item Interpret balance sheets and statements of changes in equity.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Understanding the Balance Sheet}
\begin{enumerate}
  \item What underlying equation does the balance sheet represent?
  \item 
  \item Define the following liquidity ratios:
  \begin{itemize}
    \item Current
    \item Quick
    \item Cash
  \end{itemize}
  \item Define the following solvency ratios:
  \begin{itemize}
    \item Long-term debt to equity
    \item Debt to equity
    \item Total Debt
    \item Financial Leverage
  \end{itemize}
  \item State three purposes of financial ratios.
  \item State four items that limit financial statement analysis.
\end{enumerate}

\end{document}

