\documentclass{article}

\title{Understanding the Balance Sheet}
\author{Toby Banks}
\date{October 24th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%% Components and Format of the Balance Sheet %%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Components and Format of the Balance Sheet}
The balance sheet is the usual place to start analyzing a company. It can provide information about future earnings potential. \\
\hfill \\
The balance sheet discloses what a entity owns (assets) and what it owes (liabilities) at a certain point in time. Which is why it is often referred to as a statement of financial position. \\
\hfill \\

%%ADD MORE HERE
The \textbf{report format} lists assets, liabilities and equity in a single column. \\
\hfill \\
The \textbf{account format} follows the pattern of the traditional general ledger accounts, with the assets as the left and liabilities and equity at the right of a central dividing line. \\
\hfill \\
The grouping of various classes of assets and liabilities results in a balance sheet format known as the \textbf{classified balance sheet}. \\
\hfill \\


%%%%%%%%%%%%%%%%%%%%%%%%% Uses and Analysis of the Balance Sheet %%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Uses and Analysis of the Balance Sheet}

\subsection{Common-Size Analysis of the Balance Sheet}
This technique is the process of stating all balance sheet items as a percentage of total assets and a combination of shareholders equity and liabilities. \\ 
\hfill \\
You can use this to determine things such as whether or not a company is more or less liquid from the previous year. \\
\hfill \\

\subsection{Balance Sheet Ratios}
Balance Sheet Ratios fall under the heading of \textbf{liquidity ratios} or \textbf{solvency ratios}. \\
\hfill \\
The following are some liquidity ratios: \\
\begin{itemize}
  \item Current = Current Assets / Current Liabilities
  \item Quick (acid test) = (Cash + Marketable Securities + Receivables) / Current Liabilities
  \item Cash = (Cash + Marketable Securities) / Current Liabilities
\end{itemize}
\hfill \\

The following are some solvency ratios: \\
\begin{itemize}
  \item Long-term debt to equity = Total Long-term debt / Total Equity
  \item Debt to Equity = Total Debt / Total Equity
  \item Total Debt = Total Debt / Total Assets
  \item Financial Leverage = Total Assets / Total Equity
\end{itemize}
\hfill \\

What are the purposes of financial ratios: \\
\begin{itemize}
  \item To provide insight into the microeconomic relationships within a company to help analysts project earnings and free cash flow (which is in turn used to determine entity value and credit-worthiness).
  \item To provide insight into a company's financial flexibility, which is the ability to obtain cash required to meet financial obligations or to make asset acquisitions, even if unexpected circumstances should develop.
  \item To provide a means of evaluating management's ability. Key performance ratios can serve as a quantitative measure for ranking management's ability relative to a peer group.
\end{itemize}
\hfill \\

Financial ratio analysis is limited by: \\
\begin{itemize}
  \item The use of alternative accounting methods.
  \begin{itemize}
    \item FIFO vs. Weighted vs. LIFO
    \item Cost of equity methods of accounting for unconsolidated associates
    \item Straight-line or accelerated consumption methods of deprecation
    \item Capitalized or operating lease treatment
  \end{itemize}
  \item The homogeneity of a company's operating activities. (Think G.E.)
  \item The need to determine whether the results of the ratio analysis are consistent. One set of ratios might indicate a problem while another would indicate that the problem is short term in nature.
  \item The need to use judgment. Ratios are useful, but they cannot alone be used to directly value a company or determine its credit-worthiness.
\end{itemize}
\hfill \\

\end{document}

