\documentclass{article}
\usepackage{amsmath}

\title{Common Probability Distributions}
\author{Toby Banks \\ HB-Studios}
\date{August 25th, 2009}

\begin{document}

\section{Overview}
-Four different distributions: uniform, binomial, normal, and lognormal.
-Probability distributions form the basis for hypothesis testing and regression analysis.

\section{Probability Distribution}
-A probability distribution lists all possible outcomes of a random variable and the probability that each outcome will occur. 
-A discrete probability distribution is where a random variable can take on a countable number of outcomes. i.e. Quoted Stock prices when they are quoted in ticks.
-A continuous probability distribution is one where we cannot count the outcomes. i.e. Rates or return.

A probability function specifies the probability that a random variable takes on a specific value P(X = x) (written p(x)). For a continouous function the appropriate syntax is f(x). f(x) is called the probability density function.

Key properties of a probability function:
 1. p(x) lies between 0 and 1.
 2. The sunm of the probabilitiies of p(x) over all values of X is equal to 1.

Cumulative Distribution Function (CDF):
 P(X <= x) (often denoted F(x) for both continuous and discrete random variables).

\section{Discrete Uniform Distribution}
 The probability of each outsome is the same (uniform). 

\section{Binomial Distribution}
\subsection{Bernoulli Trials}
Bernoulli trials refer to trials where there are two possible outcomes. If \emph{p} is used for the probability of success, and Y=1 denotes the outcome being a success and Y=0 a failure then:
\begin{math}p(1) = P(Y = 1) = p\end{math}
\begin{math}p(0) = P(Y = 0) = 1 - p\end{math}
Such a random variable is refered to as a Bertoulli random variable.
Trials need to be independent.
If we define X as the number of successes, then X is the binomial random variable. 
\begin{displaymath} X = Y_1 + Y_2 + ... + Y_n\end{displaymath}

X is described by n and p we write:
\begin{displaymath} X ~ B(n, p) \end{displaymath}

Which is read "X has a binomial distribution with parameters n and p".

\subsection{Binomial Probability Distribution}
Used to calculate the probability of x successes from n trials. The trials must be independent, and the probability of success across all of the trials has to be constant.
\begin{displaymath}
p\left(x\right) = {n! \over {\left(n - x\right)! x!} p^x\left(1 - p\right)^{n -x}}
\end{displaymath}

\subsubsection{Mean and Variance of Binomial Random Variables}
For \begin{math}X ~ B(n, p)\end{math} then the mean is \begin{math}np\end{math}.
Variance of a Bernoulli trial is p(1-p) (which can be calculated using \begin{math}E\left[\left(Y - EY\right)^{2}\right]\end{math}. Since trials are independent the variance of the Binomial distribution is \begin{math}np\left(1 - p\right)\end{math}.

\textbf{LOOK ELSEWHERE}Know how to construct a binomial tree and use it to describe stock movement.


\section{continuous}


\end{document}

