\documentclass{article}

\author{Toby Banks}
\date{December 3rd, 2009}

\begin{document}

\section{Organization and Functioning of Securities Markets}
\begin{itemize}
  \item Describe the characteristics of a well-functioning securities market.
  \item Distinguish between primary and secondary capital markets and explain how secondary markets support primary markets.
  \item Distinguish between call and continious markets.
  \item Compare and contrast the structural differences among national stock exchanges, regional stock exchanges, and the over-the-counter (OTC) markets.
  \item Compare and contrast major characteristics of various exchange markets, including exchange membership, types of orders, and market makers.
  \item Describe the process of selling a stock short and discuss an investor's likely motivation for selling short.
  \item Describe the process of buying a stock on margin, compute the rate of return on a margin transaction, define maintenance margin, and determine the stock price at which the investor would receive a margin call.
\end{itemize}

\begin{enumerate}
  \item What is internal efficiency of a market?
  \item What is external (or informational) efficiency of a market?
  \item What are the four characteristics that a good marker would have?
  \item What is the difference between primary and secondary markets?
  \item What are the three types of govenment bond issues and the time lengths associated with them?
  \item Contrast the three methods for issuing municipal bonds?
  \item What two types of municipal bond issues require a underwriter?
  \item The underwriting function involves three services, what are they?
  \item What are seasoned equity issues?
  \item What are IPO's?
  \item What is best-efforts basis for issuing securities?
  \item What is rule 415?
  \item What are private placements and what do they have to do with rule 415?
  \item Why are secondary markets important?
  \item What are the five classifications of markets and give some examples?
  \item What are the three old school classifications of markets?
  \item What is a pure auction market (aka. a order-driven market)?
  \item What is a dealer market (aka. a quote-driven market)?
  \item What is a call market?
  \item What is a continious market?
  \item 
\end{enumerate}

\end{document}

