\documentclass{article}

\title{Understanding the Income Statement}
\author{Toby Banks}
\date{October 12th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%%
\section{Understanding the Income Statement}
\begin{itemize}
  \item Describe the components of the income statement, and construct a income statement using the alternative presentation formats of that statement.
  \item Explain the general principles of revenue recognition and accrual accounting, demonstrate specific revenue recognition applications (including accounting for long-term contracts, installment sales, barter transactions, and gross and net reporting of revenue), and discuss the implications of revenue recognition principles for financial analysis.
  \item Discuss the general principles of expense recognition, such as the matching principle, specific expense recognition applications (including depreciation of long-term assets and inventory methods), and the implications of expense recognition principles for financial analysis.
  \item Demonstrate the appropriate method of depreciating long-term assets, accounting for inventory, or amortizing intangibles, based on the facts that might influence the decision. 
  \item Distinguish between the operating and non-operating components of the income statement.
  \item Discuss the financial reporting treatment and analysis of nonrecurring items (including discontinued operations, extraordinary items, and unusual or infrequent items) and changes in accounting standards.
  \item describe the components of earnings per share and calculate a company's earnings per share (both basic and diluted earnings per share) for both a simple and complex capital structure.
  \item Differentiate between dilutive and antidilutive securities, and discuss implications of each for the earnings per share calculation.
  \item Describe and calculate comprehensive income.
  \item State the accounting classification for items that are excluded from the income statement but affect owners' equity, and list the major types of items receiving that treatment.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Understanding the Income Statement}
\begin{enumerate}
  \item Define and present a equation for net profit, gross revenue, and operating profit.
  \item What is the difference between grouping by function and grouping by nature?
  \item What is the definition of income according to the IAS?
  \item Under the IAS what four conditions are to be met to consider a good to be sold?
  \item Under the IAS what two main conditions and four sub-conditions that are to be met to consider a service to be sold?
  \item Under the GAAP what four conditions are to be met to consider a good or service to be sold?
  \item What are the four special conditions to revenue recognition and what are some appropriate recognition strategies for each?
  \item Define expenses.
  \item What is the matching principle?
  \item Define period costs.
  \item What are three main issues with expense recognition and how do we handle those on the income statement?
  \item What is the difference between depreciation and amortization?
  \item What are the following methods of depreciation: straight line, accelerated methods, diminishing balance, and double declining.
  \item When can extraordinary items be added to the income statement?
  \item How do you handle changes in the accounting standards?
  \item How are non-operating items reported?
  \item What is the difference between simple and complex capital structure?
  \item What is diluted EPS? What is basic EPS?
  \item What is anti-diluted EPS? Give a example.
  \item How do you calculate diluted EPS when a company has convertible preferred stock outstanding?
  \item How do you calculate diluted EPS when a company has convertible debt outstanding?
  \item How do you calculate diluted EPS when a company has stock options, warrants, or their equivalents outstanding?
  \item What are two analytical tools we can use to analyze the income statement?
  \item Define the following: net profit margin, gross profit margin, operating margin, and pretax margin.
  \item What is comprehensive income?
  \item What are the four types of comprehensive income according to U.S. GAAP?
\end{enumerate}

\end{document}

