\documentclass{article}

\title{Markets for Factors of Production}
\author{Toby Banks}
\date{September 26th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%
\section{Markets for Factors of Production}
\begin{itemize}
  \item Explain why demand for the factors of production is called derived demand. Differentiate between marginal revenue and marginal revenue product (MRP). Describe how the MRP determines the demand for the labor and the wage rate.
  \item Describe the factors that cause changes in the demand for labor and the factors that determine the elasticity of demand for labor.
  \item Describe the factors determining the supply of labor, including the substitution and income effects, and discuss the factors related to changes in the supply of labor, including capital accumulation.
  \item Describe the effects on wages of labor unions and of a monopsony and explain the possible consequences for a market that offers a efficient wage.
  \item Differentiate between physical capital and financial capital and explain the relation between the demand for physical capital and the demand for financial capital.
  \item Explain the factors that influence the demand and supply of capital.
  \item Differentiate between renewable and nonrenewable natural resources and describe the supply curve for each.
  \item Differentiate between economic rent and opportunity costs.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Markets for Factors of Production}
\begin{enumerate}
  \item What is derived demand?
  \item What is marginal revenue product of labor?
  \item What are the two conditions for maximum profit? Mathematically prove they are equivalent conditions.
  \item What three factors affect the demand for labor? Give a example of each.
  \item How does the law of demand affect the quantity of labor demanded by a firm?
  \item What are the three factors that affect the elasticity of demand for labor? Give a example of each.
  \item What effect does the substitution effect and income effect on the supply of labor? How does this bend the supply curve?
  \item What trends have effected the labor market equilibrum in North America?
  \item What are the three broad objectives of a labor union?
  \item What leverage does a labor union have? What issue does this cause and how does a labor union try to get around it?
  \item What are five ways a labor union tries to increase the demand for the labor of its members?
  \item What is a monopsony? Can you draw a graph describing a monopsony labor market?
  \item What is a bilateral monopoly?
  \item What is a efficiency wage? What effects does it have?
  \item What are capital markets?
  \item What two factors determine investment and borrowing plans for firms?
  \item What three factors determine the quantity of capital saved by people?
  \item What is the Hotelling Principal? Why is it inaccurate?
  \item Compare and contrast economic rent and opportunity cost. What is its relationship to the supply curve?
  \item How does economic rent relate to taxes? Specifically, is taxing economic rent efficient?
\end{enumerate}

\end{document}

