\documentclass{article}

\title{Elasticity - Questions}
\author{Toby Banks}
\date{September 19th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%%
\section{Elasticity - Learning Outcomes}
\begin{itemize}
  \item Calculate and interpret the elasticities of demand (price elasticity, cross elasticity, and income elasticity) and the elasticity of supply and discuss the factors that influence each measure.
  \item Calculate elasticities on a straight-line demand curve, differentiate among elastic, inelastic, and unit elastic demand, and describe the relation between price elasticity of demand and total revenue.
\end{itemize}


%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Elasticity - Questions}
\begin{enumerate}
  \item What is the price elasticity of demand? How do you calculate it?
  \item define the following: perfectly inelastic demand, inelastic demand, unit elastic demand, elastic demand, perfectly elastic demand. Draw a graph for each.
  \item How is total revenue effected by the elasticity of demand?
  \item What is the total revenue test? How is it related to elasticity of demand.
  \item Can you draw two graphs detailing the relationship between elasticity and total revenue.
  \item What are the three main factors that influence the magnitude of the elasticity of demand? Detail each.
  \item What is cross elasticity of demand? How do you tell if a product is a substitute or a complement?
  \item What is income elasticity of demand?
  \item How do you define a inferior good? What is the income elasticity of demand for a inferior good?
  \item Detail five products with elastic demand and five products with inelastic demand.
  \item What is elasticity of supply? How do you calculate it?
  \item What are the two factors that influence the elasticity of supply?
  \item \textbf{BIGGIE} Create a table detailing the description of the relationship, magnitude, and meaning of the following:
  \begin{enumerate}
    \item Price Elasticity of Demand
    \item Cross Elasticity of Demand
    \item Income Elasticity of Demand
    \item Elasticity of Supply
  \end{enumerate}
\end{enumerate}

\end{document}

