\documentclass{article}

\title{Monitoring Jobs and the Price Level}
\author{Toby Banks}
\date{September 27th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%
\section{Monitoring Jobs and the Price Level}
\begin{itemize}
  \item Define an unemployed person and interpret the main labor market indicators.
  \item Define aggregate hours and real wage rates and explain their relation to gross domestic product (GDP).
  \item Explain the types of uneployment, full employment, the natural rate of unemployment, and the relation between unepmloyment and real GDP.
  \item Explain and calculate the consumer price index (CPI) and inflation rate. Describe the relation between the CPI and the inflation rate, and explain the main sources of CPI bias.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Monitoring Jobs and the Price Level}
\begin{enumerate}
  \item What are the three labor market indicators. What are the formulas for each.
  \item What are the three categories that a person who is available for work must fall into to be considered unemployed. 
  \item What does aggregate hours mean?
  \item What is the real wage rate?
  \item What are the three type of unemployment? Give a example of each.
  \item What is full employment?
  \item What is potential GDP? 
  \item What is the Consumer Price Index? What are the three broad stages to determining the CPI? 
  \item What is inflation? How do you calculate it?
  \item What are the four sources of bias in the CPI?
  \item What is the magnitude of the bias in the CPI? What results from this bias?
\end{enumerate}

\end{document}

