\documentclass{article}

\title{Monitoring Jobs and the Price Level}
\author{Toby Banks}
\date{September 27th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Jobs and Wages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Jobs and Wages}

The \textbf{working-age population} is defined as the total number of people aged 16 years and over who are not in jail, hospital, or some other form of institutional care. \\
\hfill \\
There are three indicators of the state of the labor market that are calculated by the Census Bureau:
\begin{itemize}
  \item The unemployment rate. ((Number of People Unemployed / Labor Force) * 100).
  \item The labor force participation rate. ((Labor Force / Working-age Population) * 100).
  \item The employment-to-population ratio. ((Number of People Employed / Working-age Population) * 100).
\end{itemize}
\hfill \\
The \textbf{labor force} is the sum of the employed and unemployed. \\
\hfill \\
To be counted as unemployed you must be available for work and must be one in one of the following three categories: 
\begin{enumerate}
  \item Without work bus has made specific efforts to find a job within the previous four weeks. 
  \item Waiting to be called back to a job from which he or she has been laid off. 
  \item Waiting to start a new job within 30 days.
\end{enumerate}
\hfill \\
The unemployment rate averaged 5.9 percent in the U.S. Between 1961 and 2006. It increases in recessions and decreases in expansions. \\
\hfill \\
The labor force participation rate and employment-to-population ratio have an upward trend and fluctuate with the business cycle. \\
\hfill \\
The labor force participation rate has increased for females and decreased for males. \\
\hfill \\
\textbf{Aggregate hours} are the total number of hours worked by all people employed, both full time and part time, during a year.
\hfill \\
Aggregate hours have an upward trend, and they fluctuate with the business cycle. \\
\hfill \\
\textbf{Real wage rate} is the quantity of goods and services that an hour's work can buy. \\
\hfill \\
Real hourly wage rates have been growing, but their growth rates slowed during the 1970s. \\
\hfill \\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Unemployment and Full Employment %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Unemployment and Full Employment}
People are constantly entering and leaving the state of unemployment. \\
\hfill \\
The Duration of unemployment fluctuates over the business cycle. But the demographic patterns of unemployment are constant. \\
\hfill \\
Unemployment can be frictional, structural, or cyclical. \\
\hfill \\
When all unemployment is frictional and structural, the unemployment rate equals the natural unemployment rate, the economy is at full employment, and the real GDP equals potential GDP. \\
\hfill \\
Over the business cycle, real GDP fluctuates around potential GDP and the unemployment rate fluctuates around the natural unemployment rate. \\
\hfill \\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% The Consumer Price Index %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Consumer Price Index}
The Consumer Price Index (CPI) is a measure of the average of the prices paid by urban consumers for a fixed basket of consumer goods and services. \\
\hfill \\
The CPI is defined to equal 100 for a reference base period - currently 1982-1984. \\
\hfill \\
The inflation rate is the percentage change in the CPI from one period to the next. \\
\hfill \\
Changes in the CPI probably overstate the inflation rate because of the bias that arises from new goods, quality changes, commodity substitution, and outlet substitution bias. \\
\hfill \\
The bias in the CPI distorts private contracts and increases government outlays (a Congressional Advisory Commission found the CPI overstates inflation by 1.1 percentage points per year). \\
\hfill \\
The CPI calculation has three steps:
\begin{enumerate}
  \item Find the cost of the CPI basket at base-period prices.
  \item Find the cost of the CPI basket at current-period prices.
  \item Calculate the CPI for the base period and the current period.
\end{enumerate}
\hfill \\
Inflation rate = ((CPI this year - CPI last year) / CPI last year) * 100.
\hfill \\

\end{document}

