\documentclass{article}

\author{Toby Banks}
\date{Dec 12th, 2009}

\begin{document}

\section{Working Capital Management - LOS}
\begin{itemize}
  \item Describe primary and secondary sources of liquidity and factors that influence a company's liquidity position.
  \item Compare a company's liquidity measures with those of peer companies.
  \item Evaluate overall working capital effectiveness of a company, using the operating and cash conversion cycles, and compare its effectiveness with other peer companies. 
  \item Identify and evaluate the necessary tools to use in managing a company's net daily cash position.
  \item Compute and interpret comparable yields on various securities, compare portfolio returns against a standard benchmark, and evaluate a company's short-term investment policy guidelines. 
  \item Assess the performance of a company's accounts receivable, inventory management, and accounts payable functions against historical figures and comparable peer company values.
  \item Evaluate the choices of short-term funding available to a company and recommend a financing method.
\end{itemize}

\section{Working Capital Management - Questions}
\begin{enumerate}
  \item Name the three primary sources of liquidity.
  \item Name the three secondary sources of liquidity.
  \item Name the three major drags on liquidity.
  \item Name the four major pulls on liquidity.
  \item What is the current ratio?
  \item What is the quick ratio?
  \item What is the acid-test ratio?
  \item What is the accounts receivable turnover ratio?
  \item What is the inventory turnover ratio?
  \item What is the number of days of receivables ratio?
  \item What is the number of days of inventory ratio?
  \item What is the number of days of payables ratio?
  \item What is the formula for operating cycle?
  \item What is the formula for net operating cycle?
  \item What is the purpose of managing the cash position?
%More managing the cash position questions should go here.
  \item What are the typical maturities, features, and risks for the following short-term investment instruments:
  \begin{enumerate}
    \item U.S. Treasury Bills (T-Bills)
    \item Federal Agency Securities
    \item Bank Certificates of Deposit (CDs)
    \item Banker's Acceptances (BAs)
    \item Eurodollar Time Deposits
    \item Bank Sweep Services
    \item Repurchase Agreements (Repos)
    \item Commercial Paper (CP)
    \item Mutual Funds and Money Market Mutual Funds
    \item Tax-Advantaged Securities
  \end{enumerate}
  \item What is the formula for a money-market yield?
  \item What is the formula for a bond equivalent yield?
  \item What is the formula for a discount-basis yield?
  \item What is the difference between a passive and active strategy for short-term investment?
  \item What are the following strategies: matching, mismatching, laddering. How are they different from each other?
  \item What are the five elements that make up a short-term investment policy?
% Managing Accounts Receivable
  \item What are the five goals for the accounts receivable management system?
  \item What are the four major types of credit accounts?
  \item What are five different forms of terms of credit other than cash?
  \item What are the following: EFT, ACH and Giro system?
  \item What is a lockbox system?
  \item What is the formula to calculate the float factor? What is the float factor?
  \item What is a accounts receivable aging schedule?
%Managing Inventory
  \item What are the following: transactios motive, precautionaruy stocks, stock-out losses.
  \item What is economic order quantity-reorder point (EOQ-ROP)? 
  \item What is manufacturing resource planning MRP?
%Managing Accounts Payable
  \item What are six factors that a company should consider as guidelines for effectively managing its accounts payable?
  \item What is the formula for the cost of trade credit? Why is it important?
%Managing Accounts Payable
  \item What are committed lines fo credit?
  \item What are revolvers?
  \item What are the three major objectives of a short-term borrowing strategy?
  \item What are three additional factors that borrowers should consider as part of their short-term borrowing strategies?
  \item What is a forumla to compute the cost of a credit line that requires a commitment fee?
  \item What is a forumla to compute the cost of a credit line that is all inclusive?
  \item What is a forumla to compute the cost of a credit line if dealer's fees and other fees (e.g. backup costs) are included?
\end{enumerate}

\end{document}

