\documentclass{article}

\title{Markets in Action - Questions}
\author{Toby Banks}
\date{September 19th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%%
\section{Markets in Action - Learning Outcomes}
\begin{itemize}
  \item Explain market equilibrium, distinguish between long-term and short-term effects of outside shocks, and describe the effects of rent ceilings on the existence of black markets in the housing sector and on the market's efficiency.
  \item Describe labor market equilibrium and explain the effects and inefficiencies of a minimum wage above the equilibrium wage.
  \item Explain the impact of taxes on supply, demand and market equilibrium, and describe tax incidence and its relation to demand and supply elasticity.
  \item Discuss the impact of subsides, quotas, and markets for illegal goods on demand, supply and market equilibrium.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Markets in Action - Questions}
\begin{enumerate}
  \item Describe the effects of a price ceiling on rent in the event of a earthquake that destroys a large amount of housing. Draw a graph detailing this. Also draw a graph detailing the inefficiency of a rent ceiling, this graph should include sections for consumer surplus, producer surplus, deadweight loss, and potential loss from housing search.
  \item What is search activity?
  \item What is a black market?
  \item What effect does the minimum wage have on the labor market. What effect does the living wage have?
  \item Tax incidence is the division of the burden of a tax between the buyer and the seller. Why is it not obvious who pays taxes? 
  \item What is the effect of the elasticity of demand on who pays the brunt of the taxes (the buyer or seller).
  \item What effect does taxes have on the efficiency of the marketplace?
  \item What is a subsidy? What effect does it have on production and consumption of goods and services?
  \item What is a production quota? What effect does it have on production and consumption of goods and services?
  \item What effect does having penalties on buyers and sellers of illegal goods or services have?
\end{enumerate}

\end{document}

