\documentclass{article}

\title{Financial Analysis Techniques}
\author{Toby Banks}
\date{October 31st, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%%
\section{Financial Analysis Techniques}
\begin{itemize}
  \item Evaluate and compare companies using ratio analysis, common-size financial statements, and charts in financial analysis.
  \item Describe the limitations of ratio analysis.
  \item Describe the various techniques of common-size analysis and interpret the results of such analysis.
  \item Calculate, classify, and interpret activity, liquidity, solvency profitability, and valuation ratios.
  \item Demonstrate how ratios are related and how to evaluate a company using a combination of different ratios.
  \item Demonstrate the application of and interpret changes in the component parts of the DuPont analysis (the decomposition of return on equity).
  \item Calculate and interpret the ratios used in equity analysis, credit analysis, and segment analysis. 
  \item Describe how ratio analysis and other techniques can be used to model and forecast earnings.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Financial Analysis Techniques}
\begin{enumerate}
  \item What five questions should you ask yourself before beginning a financial analysis?
  \item Detail an overview of the financial statement analysis framework. (LONG ANSWER)
  \item What are four main limitations of ratio analysis? Go into details for each.
  \item What is common size analysis? How does it differ for the income and balance sheets?
  \item When would one use vertical stacks, pie charts or line charts?
  \item What is regression analysis?
  \item What are the five categories of financial ratios. List all of the common studied ratios for each.
  \item Define and describe the use of the following activity ratios:
  \begin{enumerate}
    \item Inventory Turnover
    \item Days of Inventory on Hand (DOH)
    \item Receivables Turnover
    \item Days of Sales Outstanding (DSO)
    \item Payables Turnover
    \item Number of Days of Payables
    \item Working Capital Turnover
    \item Fixed Asset Turnover
    \item Total Asset Turnover
  \end{enumerate}
  \item Define and describe the uses of the following liquidity ratios.
  \begin{enumerate}
    \item Current Ratio
    \item Quick Ratio
    \item Cash Ratio
    \item Defensive Interval Ratio
  \end{enumerate}
  \item Explain the cash conversion cycle.
  \item Define and describe the uses of the following solvency debt ratios.
  \begin{enumerate}
    \item Debt-To-Assets Ratio (Total Debt Ratio)
    \item Debt-To-Capital Ratio
    \item Debt-To-Equity Ratio
    \item Financial Leverage Ratio
  \end{enumerate}
  \item Define and describe the uses of the following solvency coverage ratios.
  \begin{enumerate}
    \item Interest Coverage
    \item Fixed Charge Coverage
  \end{enumerate}
  \item Define and describe the uses of the following return on sales profitability.
  \begin{enumerate}
    \item Gross Profit Margin
    \item Operating Profit Margin
    \item Pretax Profit Margin
    \item Net Profit Margin
  \end{enumerate}
  \item Define and describe the uses of the following return on investment profitability.
  \begin{enumerate}
    \item Operating ROA
    \item ROA (return on assets)
    \item Return on Total Capital
    \item ROE
    \item Return on Common Equity
  \end{enumerate}
  \item When decomposing ROE what are the two different mutually exclusive ways of representing the interest burden.
  \item Draw a full ROE tree. 
  \item What does each of the elements in a ROE tree mean?
  \item What are five general steps when evaluating a security for inclusion in a portfolio?
  \item Define the following valuation ratios and what they mean:
  \begin{enumerate}
    \item P/E
    \item P/CF
    \item P/S
    \item P/BV
  \end{enumerate}
  \item Define the following as ratios:
  \begin{enumerate}
    \item Basic EPS
    \item Diluted EPS
    \item Cash Flow Per Share
    \item EBITDA per share
    \item Dividends per share
  \end{enumerate}
  \item Define the following dividend related quantiles:
  \begin{enumerate}
    \item Dividend payout ratio
    \item Retention rate (b)
    \item Sustainable Growth Rate
  \end{enumerate}
  \item Detail why dividend related quantiles are used and why?
  \item Define the following credit analysis ratios:
  \begin{enumerate}
    \item EBIT Interest Coverage
    \item EBITDA Interest Coverage 
    \item Funds from operations to total debt
    \item Free operating cash flow to total debt
    \item Total Debt to EBITDA
    \item Return on Capital
    \item Total Debt to Total Debt plus Equity
  \end{enumerate}
  \item What nine items must be reported for a primary business segment?
  \item What three items must be reported for a secondary business segment?
  \item What other five items must be disclosed regarding business/geographic segments?
  \item Define the following four segment ratios:
  \begin{enumerate}
    \item Segment Margin
    \item Segment Turnover
    \item Segment ROA
    \item Segment Debt Ratio
  \end{enumerate}
  \item What are three techniques used in model building and forecasting?
\end{enumerate}

\end{document}

