\documentclass{article}

\title{Monopoly - Questions}
\author{Toby Banks}
\date{September 20th, 2009}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%% Learning Outcomes %%%%%%%%%%%%%%%%%%%%%%%
\section{Monopoly - Learning Outcomes}
\begin{itemize}
  \item Describe the characteristics of a monopoly, including factors that allow a monopoly to arise and monopoly price-setting strategies.
  \item Explain the relation between price, marginal revenue, and elasticity for a monopoly and determine a monopoly's profit-maximizing price and quantity.
  \item Explain price discrimination and why perfect price discrimination is efficient.
  \item Explain how consumer and producer surplus are redistributed in a monopoly.
  \item Explain the potential gains from monopoly and the regulation of a natural monopoly.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%% Questions %%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Monopoly - Questions}
\begin{enumerate}
  \item What is market power?
  \item What are the two main features of a monopoly?
  \item What are two different barriers to entry for firms looking to compete with a monopoly?
  \item What are the two different price setting strategies that monopolys employ? Go into detail about each.
  \item What is economic rent?
  \item What is rent-seeking?
  \item What is rent-seeking equilibrium?
  \item What two factors enable a monopoly to practice price discrimination?
  \item What are two broad ways that monopolies discriminate to try to extract consumer surplus?
  \item What is perfect price discrimination? Who benefits? Is it efficient? 
  \item What are the two major types of advantages that monopolys bring to the table? Detail each.
  \item What is the marginal cost pricing rule? What can monopolies do to cover its costs if this rule is in effect?
  \item What is average cost pricing rule? 
\end{enumerate}

\end{document}

