\documentclass{article}

\author{Toby Banks}
\date{Dec 3rd, 2009}

\begin{document}
\textbf{Common Mistakes}

\textbf{Documentation}
\begin{itemize}
  \item Do all new functions and classes have headers?
  \item Are all comments consistent with the code?
  \item Does all code that is commented/preprocessored out have an explanation?
  \item Are all parameters and variables properly described, either by their names, or by comments?
\end{itemize}

Perforce
\begin{itemize}
  \item Does the changelist description properly describe all the changes?
  \item Are the changes made against the latest version of the file?
  \item Are all the necessary data files in the changelist?
\end{itemize}

\textbf{Coding Standard and Readability}
\begin{itemize}
  \item Do all class/function/variable names conform to the HB Coding Standards?
  \item Do all class/function/variable names have clearly unique, understandable names?
  \item Is the code properly formatted, in terms of identation, use of whitespace, etc?
  \item Do sets of functions that process similar parameters have their parameters in a consistent order?
  \item Has spaghetti code been avoided by appropriate use of functions?
  \item Are all special values represented by const values or enums?
\end{itemize}


\textbf{Low Level Concepts}
\begin{itemize}
  \item Are all variables properly initialized? Do constants have the proper suffix (i.e. float l\_fBlah = 1.5f;)?
  \item Are there any compiler errors or warnings? Have you checked platforms/builds?
  \item Are all "off-by-one" errors avoided?
  \item Are all divide by zero errors avoided?
  \item Has all allocated memory been freed in the appropriate place?
  \item Has all floating point comparison for equality been removed?
  \item Are if/else/switch statements set up properly (i.e. common cases tested first, handling all possible values, etc)?
  \item Has all c-style casting ((int)l\_fBlah) been replaces with c++ style (static\_cast<int>(l\_fBlah))?
  \item Are boolean operators (i.e. ==, \&\&, ||, etc) present as intended (i.e. not =, \&, or |)?
  \item Are switch cases properly seperated by break; statements? Do all switches handle a default case?
  \item Have all gotos been removed?
  \item Are all one-line loop/conditional statements re-formatted to be multiple lines?
  \item Are all statement blocks for loops/conditionals wrapped in curly braces?
	• Do all function calls have their parameters passed in the proper order?
\end{itemize}


\textbf{High Level Concepts}
\begin{itemize}
  \item Have all blocks of repeated code been condensed into a single function?
  \item Is all platform (pc, ps2, etc) and build (debug, final, etc) specific code wrapped in the proper preprocessor lines?
  \item Are non-primitive function parameters passed by reference instead of by value?
  \item Is the const keyword used where it should be?
  \item Does the changed code do what it is supposed do?
  \item Has all code that is no longer used been deleted, or at least commented out?
  \item Are member variables initialized in an initializer list where possible?
  \item Are member variables and methods as restricted as possible?
  \item Are all members common to derived classes moved to the base class?
\end{itemize}

\textbf{Possible Trouble Spots}
\begin{itemize}
  \item Has a new class been declared?
  \item Is the functionality only valid in certain builds? (i.e. debug only, North America only, etc)
  \item Is there some functionality to the feature that is to be done later?
  \item Were global variables/data added?
  \item Do any of the files in the changelist have to be integrated?
  \item Has in	teraction with data files been added or changed? Have any new data/art files been added to the project?
\end{itemize}

\end{document}
